---
title: "Chemoproteomics MSstats comparison"
author: "Sarah Szvetecz"
date: "2023-12-01"
output: html_document
---
Load packages
```{r}
library(MSstats)
library(readr)
library(dplyr)
library(ggplot2)
library(reshape)

```

Notes: 
each compound processed as it's own dataset with MaxQuant 
- data in folders: Capmatinib, Lapatinib, Midostaurin, Rabusertib, and Tanzisertib 

1) Create annotation files 
  - reference paper
  - maps ms stats run ... raw file, condition, bioreplciate 
  - msstats4.0, group comparison, time series, 

2) Run MaxQtoMSstatsFormat converter on each dataset
  - MaxQtoMSstatsFormat requires evidence and proteinGroups (and annotation?) txt files
  - manually make annotation file in code 
  
Look at data
```{r}
summary <- read.table( "/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/summary.txt",header = TRUE, fill = TRUE, sep = '\t')
summary <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/summary.txt")

head(summary)

#proteinGroups
proteinGroups <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/proteinGroups.txt")
colnames(proteinGroups)
head(proteinGroups)
dim(proteinGroups) #3252 x 208
View(proteinGroups)
#evidence

evidence <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/evidence.txt")

head(evidence)
dim(evidence) #298256  x   67
#take raw files 
length(unique(evidence$`Raw file`)) #25 runs ; 15 DMSO controls ; 



#normalization using PDPD and DMSO
length(unique(evidence$Proteins))
```

Run for each compound
```{r}
#set path
path<-"/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/"
compounds<-c("Capmatinib","Lapatinib","Midostaurin","Rabusertib","Tanzisertib")

#create input files 
input_files<-c()
for(compound in compounds){
temp_path_evidence<-paste(path, compound, "/txt/evidence.txt",sep="")
temp_path_pg<-paste(path, compound, "/txt/proteinGroups.txt",sep="")

#read txt files 
evidence<-read_tsv(temp_path_evidence)
proteinGroups<-read_tsv(temp_path_pg)

annot = evidence %>% distinct(`Raw file`, Experiment)
colnames(annot)[2]<-"Condition"
annot$BioReplicate = 1
input = MaxQtoMSstatsFormat(evidence, annot, proteinGroups,)

input_files[[compound]]<-input

}

?MaxQtoMSstatsFormat


input_files[[1]]
compound<-compounds[1]
cap_evidence<-paste(path, compound, "/txt/evidence.txt",sep="")
View(read_tsv(cap_evidence))
cap_evidence<-read_tsv(cap_evidence)
```

Check variable length (# of proteins and # of peptides) -- 
```{r}
length(unique(input_files[[1]]$ProteinName)) #2538
length(unique(input_files[[1]]$PeptideSequence)) #19799

length(unique(input_files[[2]]$ProteinName)) #2697
length(unique(input_files[[2]]$PeptideSequence)) #21995
unique(input_files[[2]]$Condition)
table(input_files[[2]]$Condition)
```

Focus on Capmatinib
Clean data --- remove all run controls c1:c15 
```{r}
test_drug1<-input_files[[1]]
unique(test_drug1$Condition)
control_remove<-paste("c",1:15,sep="")

#remove run controls (c1:c15) ... we don't use these in the analysis 
test_drug1_clean<-subset(test_drug1 ,!Condition %in% control_remove)
unique(test_drug1_clean$Condition)

```


Check missing values before dataProccess step 
```{r}
#MaxQuant output --- 
cap_NAsMaxQuant<-cap_evidence %>% filter(is.na(`Max intensity m/z 0`)) 
length(unique(cap_NAsMaxQuant$`Protein names`))
#MSstats format --- 
cap_NAs<-test_drug1_clean %>% filter(is.na(Intensity)) 
length(unique(cap_NAs$ProteinName)) #2489
table(cap_NAs$Condition)

```



Calculate protein level values (compute equalize medians normalization for comparison)

Note: When looking at the data noticed that values are missing at 30000nM (highest concentration)
  - Checks show that nothing is being filtered out for being "too low" in the dataset ... the values are just not            measured (NA values)
  
```{r}
#default , proper way to run
QuantData<- dataProcess(test_drug1_clean,normalization = FALSE,MBimpute = FALSE) 

#run to check if proteins are being filtered out 
QuantData_nofilter<- dataProcess(test_drug1_clean,normalization = FALSE,MBimpute = FALSE,maxQuantileforCensored = NULL) 
#not being filtered out, just not measured 

dim(QuantData$ProteinLevelData)
dim(QuantData_nofilter$ProteinLevelData)

sum(is.na(QuantData$ProteinLevelData$LogIntensities))
sum(is.na(QuantData_nofilter$ProteinLevelData$LogIntensities))

#0 rows shows that no values are being censored
QuantData_nofilter$FeatureLevelData %>% filter(!is.na(INTENSITY) & censored==TRUE) %>% head()
QuantData$FeatureLevelData %>% filter(!is.na(INTENSITY) & censored==TRUE) %>% head()

?dataProcess
#length(unique(QuantData$ProteinLevelData$Protein))

QuantData_norm <- dataProcess(test_drug1_clean,normalization = "EQUALIZEMEDIANS",MBimpute = FALSE) 
#length(unique(QuantData$ProteinLevelData$Protein))
```
Check missing values after dataProccess step 
```{r}
#QuantData --- 
#proteinlevel doesn't show NAs ... all in feature level data 
QuantData$ProteinLevelData %>% filter(!is.na(LogIntensities) & Protein == "A0AV96") 
QuantData$ProteinLevelData %>% filter( Protein == "A0AV96") 


#count number of missing values 
sub_protein<-QuantData$FeatureLevelData %>% filter(is.na(INTENSITY) & PROTEIN == "P55263") 
nrow(sub_protein)

#loop through all --- 
protien_names<-unique(QuantData$ProteinLevelData$Protein)
missing_protien_list<-data.frame(matrix(nrow=length(protien_names),ncol=3))
colnames(missing_protien_list)<-c("Protein","NumberTotal","NumberMissing")

#how much is missing in protein level data ? 
for (i in 1:length(protien_names)){
sub_protein<-QuantData$ProteinLevelData %>% filter(Protein == protien_names[i]) 
missing_protien_list$Protein[i]<-as.character(protien_names[i])
missing_protien_list$NumberTotal[i]<-nrow(sub_protein)
missing_protien_list$NumberMissing[i]<- 10 - nrow(sub_protein)
}

# make histogram 
ggplot(missing_protien_list, aes(x=NumberMissing)) + 
  geom_histogram(color="black", fill="white")

# make histogram 
ggplot(missing_protien_list, aes(x=NumberTotal)) + 
  geom_histogram(color="black", fill="white")


 
#calculate number of proteins with at least 1 missing group 
missing_protien_list %>% filter( NumberTotal != 10) 
993 / length(protien_names) #2190


#set up my data 
group_missing<-as.data.frame(c("DMSO","PDPD","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM","DMSO","PDPD","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM"))
group_missing$Missing<-c(rep("No",10),rep("Yes",10))
colnames(group_missing)[1]<-"Group"

for (i in 1:10){
temp_group<-QuantData$ProteinLevelData %>% filter(GROUP == group_missing$Group[i]) 
group_missing$count[i]<-nrow(temp_group)
group_missing$count[i+10]<-length(protien_names) - nrow(temp_group)
}

#percent DMSO missing --- 
group_missing %>% filter(Group == "DMSO") 
366 /length(protien_names)

#percent 30000uM missing --- 
group_missing %>% filter(Group == "30000nM") 
292 /length(protien_names)


# creating plot using the above data
order<-c("DMSO","PDPD","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM")
ggplot(group_missing, aes(factor(Group,level = order), count, fill = Missing)) +
  geom_bar(stat="identity", position = "dodge") + 
  labs(title="Number of missing measurements per group") +xlab("Dose")


```


```{r}
#set up loop
for (i in 1:length(protien_names)){
sub_protein<-QuantData$FeatureLevelData %>% filter(is.na(INTENSITY) & PROTEIN == protien_names[i]) 
missing_protien_list$Protein[i]<-as.character(protien_names[i])
missing_protien_list$NumberMissing[i]<-nrow(sub_protein)
}

# make histogram 
ggplot(missing_protien_list, aes(x=NumberMissing)) + 
  geom_histogram(color="black", fill="white")

#subset to 10 or less 
ggplot(missing_protien_list %>% filter(NumberMissing < 11), aes(x=NumberMissing)) + 
  geom_histogram(color="black", fill="white")


#########################################
## Repeat at feature level 

#number missing for each feature 
peptide_names<-unique(QuantData$FeatureLevelData$PEPTIDE)
missing_feature_list<-data.frame(matrix(nrow=length(peptide_names),ncol=2))
colnames(missing_feature_list)<-c("Peptide","NumberMissing")

#set up loop
for (i in 1:length(peptide_names)){
sub_peptide<-QuantData$FeatureLevelData %>% filter(is.na(INTENSITY) & PEPTIDE == peptide_names[i])
missing_feature_list$Peptide[i]<-as.character(peptide_names[i])
missing_feature_list$NumberMissing[i]<-nrow(sub_peptide)
}


# make histogram 
ggplot(missing_feature_list, aes(x=NumberMissing)) + 
  geom_histogram(color="black", fill="white")


```


Calculate Percent depletion for each protein 
```{r}
sub_DMSO<-QuantData$ProteinLevelData %>% filter(GROUP == "DMSO") 
protein_sub<-unique(sub_DMSO$Protein)

#test 
sub_DMSO_temp<-QuantData$ProteinLevelData %>% filter(GROUP == "DMSO" & Protein == protein_sub[1]) 
sub_other_temp<-QuantData$ProteinLevelData %>% filter( Protein == protein_sub[1]) 

temp_data<-as.data.frame(sub_other_temp$GROUP)
temp_data$relative_intensity<- sub_other_temp$LogIntensities / sub_DMSO_temp$LogIntensities
temp_data$Protein<-as.character(protein_sub[1])

temp_data_update<-rbind(temp_data_update,temp_data)


#loop 
temp_data_update<-as.data.frame(matrix(ncol=3))
colnames(temp_data_update)<-c("Group","relative_intensity","Protein")
for (i in 1: length(protein_sub)){
sub_DMSO_temp<-QuantData$ProteinLevelData %>% filter(GROUP == "DMSO" & Protein == protein_sub[i]) 
sub_other_temp<-QuantData$ProteinLevelData %>% filter( Protein == protein_sub[i]) 

temp_data<-as.data.frame(sub_other_temp$GROUP)
temp_data$relative_intensity<- sub_other_temp$LogIntensities / sub_DMSO_temp$LogIntensities
temp_data$Protein<-as.character(protein_sub[i])
colnames(temp_data)[1]<-"Group"

temp_data_update<-rbind(temp_data_update,temp_data)
}


View(temp_data_update)
```


Variance calculation 
```{r}
protien_names<-unique(QuantData$ProteinLevelData$Protein)

protein_var_list<-data.frame(matrix(nrow=length(protien_names),ncol=2))
#protein_var_list<-as.data.frame(protein_var_list)
colnames(protein_var_list)<-c("Protein","Variance")

#set up loop
for (i in 1:length(protien_names)){
sub_protein<-QuantData$ProteinLevelData[QuantData$ProteinLevelData$GROUP %in% c("DMSO","PDPD")& QuantData$ProteinLevelData$Protein %in% protien_names[i],]

protein_var_list$Protein[i]<-as.character(protien_names[i])
protein_var_list$Variance[i]<-var(sub_protein$LogIntensities)
}

#protein_var_list$Protein[1]<-as.character(protien_names[1])
protein_var_list
summary(protein_var_list$Variance)
546/nrow(protein_var_list)
```




Set up contrast matrix --- 
Comparison with no groupings (no reps)
```{r}
## set up comparison DMSO vs highest dose , and  DMSO  vs PDPD
comparison1<-matrix(c(0,1,0,0,0,0,0,0,0,-1),nrow=1)
comparison2<-matrix(c(0,0,0,0,0,0,0,0,1,-1),nrow=1)
comparison<-rbind(comparison1,comparison2)
row.names(comparison)<-c("DMSO-30000nM","DMSO-PDPD")
colnames(comparison)<-unique(QuantData$ProteinLevelData$GROUP)


testResultMultiComparisons<-groupComparison(contrast.matrix=comparison,data=QuantData)
View(testResultMultiComparisons$ComparisonResult) #can only look at logFC when 1 rep in each group 
```
check DMSO vs PDPD
```{r}
summary(testResultMultiComparisons$ComparisonResult[testResultMultiComparisons$ComparisonResult$Label == "DMSO-PDPD","log2FC"])

hist(testResultMultiComparisons$ComparisonResult[testResultMultiComparisons$ComparisonResult$Label == "DMSO-PDPD","log2FC"])
```
check DMSO vs 30000nM
```{r}
summary(testResultMultiComparisons$ComparisonResult[testResultMultiComparisons$ComparisonResult$Label == "DMSO-30000nM","log2FC"])


hist(testResultMultiComparisons$ComparisonResult[testResultMultiComparisons$ComparisonResult$Label == "DMSO-30000nM","log2FC"])

```


Set up contrast matrix --- 
rename columns to be same to group conditions together
*subject = rep 
```{r}
#create new QuantData for grouping samples together 
QuantData_groups<-QuantData

#rename in FeatureData
#PDPD to DMSO ---
levels(QuantData_groups$FeatureLevelData$SUBJECT)<-c(1,2)

#define as replicate
QuantData_groups$FeatureLevelData[which(QuantData_groups$FeatureLevelData$GROUP=="PDPD"),]$SUBJECT <- 2
table(QuantData_groups$FeatureLevelData$SUBJECT)

QuantData_groups$FeatureLevelData[which(QuantData_groups$FeatureLevelData$GROUP=="PDPD"),]$GROUP <- "DMSO"
table(QuantData_groups$FeatureLevelData$GROUP)

#3000nM to 30000nM ---
QuantData_groups$FeatureLevelData[which(QuantData_groups$FeatureLevelData$GROUP=="3000nM"),]$SUBJECT <- 2
table(QuantData_groups$FeatureLevelData$SUBJECT)

QuantData_groups$FeatureLevelData[which(QuantData_groups$FeatureLevelData$GROUP=="3000nM"),]$GROUP <- "30000nM"
table(QuantData_groups$FeatureLevelData$GROUP)


#rename in ProteinData
QuantData_groups$ProteinLevelData[which(QuantData_groups$ProteinLevelData$GROUP=="PDPD"),]$SUBJECT <- 2
table(QuantData_groups$ProteinLevelData$SUBJECT)

QuantData_groups$ProteinLevelData[which(QuantData_groups$ProteinLevelData$GROUP=="PDPD"),]$GROUP <- "DMSO"
table(QuantData_groups$ProteinLevelData$GROUP)

#3000nM to 30000nM ---
QuantData_groups$ProteinLevelData[which(QuantData_groups$ProteinLevelData$GROUP=="3000nM"),]$SUBJECT <- 2
table(QuantData_groups$ProteinLevelData$SUBJECT)

QuantData_groups$ProteinLevelData[which(QuantData_groups$ProteinLevelData$GROUP=="3000nM"),]$GROUP <- "30000nM"
table(QuantData_groups$ProteinLevelData$GROUP)


## set up comparison DMSO vs highest dose , and top 2 high vs top 2 low
comparison1<-matrix(c(0,1,0,-1,0,0,0,0),nrow=1)
#comparison2<-matrix(c(0,1,0,0,0,0,0,0,0,-1),nrow=1)
#comparison3<-matrix(c(-1,0,0,0,0,0,0,0,1,0),nrow=1)
#comparison<-rbind(comparison1,comparison2)
#row.names(comparison)<-c("DMSO_PDPD-30000nM_3000nM","DMSO-30000nM")
#colnames(comparison)<-unique(QuantData_groups$ProteinLevelData$GROUP)
row.names(comparison1)<-c("DMSO_PDPD-30000nM_3000nM")
comparison1<-data.frame(comparison1)
colnames(comparison1)<-unique(QuantData_groups$ProteinLevelData$GROUP)


testResultMultiComparisons_groups<-groupComparison(contrast.matrix=comparison1,data=QuantData_groups)

View(testResultMultiComparisons_groups$ComparisonResult)
```


Next test 
```{r}
#create test comparison ... less missing data? No difference here, still NAs
comparisontest<-matrix(c(0,0,0,0,0,0,0,0,1,-1),nrow=1)
#comparison<-rbind(comparison1,comparison2)
row.names(comparisontest)<-c("DMSO-30nM")
colnames(comparisontest)<-unique(QuantData$ProteinLevelData$GROUP)

testResultMultiComparisons.test<-groupComparison(contrast.matrix=comparisontest,data=QuantData)

testResultMultiComparisons;testResultMultiComparisons.test
```

Comparison using normalized data ---
```{r}
QuantData_norm

## set up comparison DMSO vs highest dose , and top 2 high vs top 2 low
comparison1<-matrix(c(0,1,0,-1,0,0,1,0,0,-1),nrow=1)
comparison2<-matrix(c(0,1,0,0,0,0,0,0,0,-1),nrow=1)
#comparison3<-matrix(c(-1,0,0,0,0,0,0,0,1,0),nrow=1)
comparison_temp<-rbind(comparison1,comparison2)
row.names(comparison_temp)<-c("DMSO_PDPD-30000nM_3000nM","DMSO-30000nM")
colnames(comparison_temp)<-unique(QuantData_norm$ProteinLevelData$GROUP)

testResultMultiComparisons_temp<-groupComparison(contrast.matrix=comparison_temp,data=QuantData_norm)

```


Visualize the log intensisties for a single protein ---- 
```{r}
test_CDKL<-subset(QuantData[[2]],Protein %in% c("O76039"))

test_Q6ZTR7<-subset(QuantData[[2]],Protein %in% c("Q6ZTR7"))

test_O15197<-subset(QuantData[[2]],Protein %in% c("O15197"))

test_Q9UI12<-subset(QuantData[[2]],Protein %in% c("Q9UI12"))

#define order
order<-c("DMSO","PDPD","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM")

test_CDKL<-test_CDKL[order(factor(test_CDKL$GROUP, levels = order)),]

test_CDK4<-test_CDK4[order(factor(test_CDK4$GROUP, levels = order)),]

#remove PDPD
`%notin%` <- Negate(`%in%`)
test_CDK4<-subset(test_CDK4,GROUP %notin% c("PDPD"))

#visualize points for 1 protein 
ggplot(test_Q9UI12,aes(x=factor(GROUP, level = order),y=LogIntensities))+geom_boxplot() + ggtitle(paste0("Test data for Protein: ", test_Q9UI12$Protein[1]))+xlab("Drug Concentration")+ylab("Log Intensity")

ggplot(test_Q9H0U3,aes(x=factor(GROUP, level = order),y=LogIntensities))+geom_boxplot() + ggtitle(paste0("Test data for Protein: MAGT1"))+xlab("Drug Concentration")+ylab("Log Intensity")

ggplot(test_Q9H0U3,aes(x=factor(GROUP, level = order),y=LogIntensities/20.33898))+geom_boxplot() + ggtitle(paste0("Test data for Protein: ", test_Q9H0U3$Protein[1]))+xlab("Drug Concentration")+ylab("Log Intensity relative to DMSO")
```


```{r}
add_ratios <- function(df, cols, ratio_cols, ref_cols) {
  # Calculate ratios of cols / mean(ref_cols)
  df[ratio_cols] <- df[cols] / rowMeans(df[ref_cols, drop = FALSE])
  
  # Replace infinity values with NA
  df[ratio_cols][is.infinite(df[ratio_cols])] <- NA
  
  return(df)
}
```

Only works for single comparison
```{r}
# Function: groupComparisonPlots
# visualization for testing results

# Volcano plot with FDR cutoff = 0.05 and no FC cutoff
groupComparisonPlots(data=testResultMultiComparisons_groups$ComparisonResult %>% filter( !is.na(pvalue)),type="VolcanoPlot",logBase.pvalue=2,sig = 0.5,address="Ex1_")



?groupComparisonPlots
# Volcano plot with FDR cutoff = 0.05, FC cutoff = 70, upper y-axis limit = 100, and no protein name displayed
# FCcutoff=70 is for demonstration purpose
groupComparisonPlots(data=testResultMultiComparisons$ComparisonResult,type="VolcanoPlot",FCcutoff=70, logBase.pvalue=2, ylimUp=100, ProteinName=FALSE,address="Ex2_")
```

Visualizations 
```{r}
# 1. Profile plot
dataProcessPlots(data=QuantData,type="ProfilePlot",width=20, height=10)

# 2. Quality control plot 
dataProcessPlots(data=QuantData,type="QCPlot")	

# 3. Quantification plot for conditions
dataProcessPlots(data=QuantData,type="ConditionPlot")

```

These won't work when doing a single comparison:
```{r}
# Heatmap with FDR cutoff = 0.05
groupComparisonPlots(data=testResultMultiComparisons$ComparisonResult,type="Heatmap", logBase.pvalue=2, address="Ex1_")

# Heatmap with FDR cutoff = 0.05 and FC cutoff = 70
# FCcutoff=70 is for demonstration purpose
groupComparisonPlots(data=testResultMultiComparisons$ComparisonResult,type="Heatmap",FCcutoff=70, logBase.pvalue=2, address="Ex2_")

# Comparison Plot
groupComparisonPlots(data=testResultMultiComparisons$ComparisonResult,type="ComparisonPlot",address="Ex1_")

# Comparison Plot
groupComparisonPlots(data=testResultMultiComparisons$ComparisonResult,type="ComparisonPlot",ylimUp=8,ylimDown=-1,address="Ex2_")
```

Extra (haven't run) ----- 

Sample size & power calculation 
```{r}
## Calculate sample size for future experiments:
#(1) Minimal number of biological replicates per condition
designSampleSize(data=testResultMultiComparisons$FittedModel, numSample=TRUE,
                 desiredFC=c(1.25,1.75), FDR=0.05, power=0.8)

#(2) Power calculation
designSampleSize(data=testResultMultiComparisons$FittedModel, numSample=2,
                 desiredFC=c(1.25,1.75), FDR=0.05, power=TRUE)

```


