---
title: "Data Processing / Explore"
author: "Sarah Szvetecz"
date: "2023-10-03"
output: html_document
---

Load packages
```{r}
library(MSstats)
library(readr)
library(dplyr)
library(ggplot2)
library(reshape)

#packages for curve fitting 
library(drc)
library(plotrix)
library(dr4pl)
```

Notes: 
each compound processed as it's own dataset with MaxQuant 
- data in folders: Capmatinib, Lapatinib, Midostaurin, Rabusertib, and Tanzisertib 

1) Create annotation files 
  - reference paper
  - maps ms stats run ... raw file, condition, bioreplciate 
  - msstats4.0, group comparison, time series, 

2) Run MaxQtoMSstatsFormat converter on each dataset
  - MaxQtoMSstatsFormat requires evidence and proteinGroups (and annotation?) txt files
  
Look at data
```{r}
summary <- read.table( "/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/summary.txt",header = TRUE, fill = TRUE, sep = '\t')
summary <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/summary.txt")

head(summary)

#proteinGroups
proteinGroups <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/proteinGroups.txt")
colnames(proteinGroups)
head(proteinGroups)
dim(proteinGroups) #3252 x 208

#evidence

evidence <- read_tsv("/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/Capmatinib/txt/evidence.txt")

head(evidence)
dim(evidence) #298256  x   67
#take raw files 
length(unique(evidence$`Raw file`)) #25 runs ; 15 DMSO controls ; 

annot


#normalization using PDPD and DMSO
length(unique(evidence$Proteins))
```

Run for each compound
```{r}
#set path
path<-"/Users/sarahszvetecz/Desktop/R/Chemoproteomics/klaeger_etal_science/"
compounds<-c("Capmatinib","Lapatinib","Midostaurin","Rabusertib","Tanzisertib")

#create input files 
input_files<-c()
for(compound in compounds){
temp_path_evidence<-paste(path, compound, "/txt/evidence.txt",sep="")
temp_path_pg<-paste(path, compound, "/txt/proteinGroups.txt",sep="")

#read txt files 
evidence<-read_tsv(temp_path_evidence)
proteinGroups<-read_tsv(temp_path_pg)

annot = evidence %>% distinct(`Raw file`, Experiment)
colnames(annot)[2]<-"Condition"
annot$BioReplicate = 1
input = MaxQtoMSstatsFormat(evidence, annot, proteinGroups)

input_files[[compound]]<-input

}


input_files[[1]]
compound<-compounds[1]
cap_evidence<-paste(path, compound, "/txt/evidence.txt",sep="")
View(read_tsv(cap_evidence))
```

Check  variables  
```{r}
length(unique(input_files[[1]]$ProteinName)) #2538
length(unique(input_files[[1]]$PeptideSequence)) #19799

length(unique(input_files[[2]]$ProteinName)) #2697
length(unique(input_files[[2]]$PeptideSequence)) #21995
unique(input_files[[2]]$Condition)
table(input_files[[2]]$Condition)
```

Convert to protein level values for Capmatinib
```{r}
test_drug1<-input_files[[1]]
unique(test_drug1$Condition)
control_remove<-paste("c",1:15,sep="")

#remove run controls (c1:c15)
test_drug1_clean<-subset(test_drug1 ,!Condition %in% control_remove)
unique(test_drug1_clean$Condition)
QuantData1 <- dataProcess(test_drug1_clean,normalization = FALSE,MBimpute = FALSE) 
length(unique(QuantData1$ProteinLevelData$Protein))

#save Capmatinib data
#write.csv(QuantData1[[2]], "Capmatinib_protien_intensities.csv",row.names=F)
```

Visualize the data 
```{r}
test_CDKL<-subset(QuantData1[[2]],Protein %in% c("O76039"))

test_CDK4<-subset(QuantData1[[2]],Protein %in% c("P11802"))

test_Q9H0U3<-subset(QuantData[[2]],Protein %in% c("Q9H0U3"))


#define order
order<-c("DMSO","PDPD","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM")

test_CDKL<-test_CDKL[order(factor(test_CDKL$GROUP, levels = order)),]

test_CDK4<-test_CDK4[order(factor(test_CDK4$GROUP, levels = order)),]

#remove PDPD
`%notin%` <- Negate(`%in%`)
test_CDK4<-subset(test_CDK4,GROUP %notin% c("PDPD"))

#visualize points for 1 protein 
ggplot(test_CDK4,aes(x=factor(GROUP, level = order),y=LogIntensities))+geom_boxplot() + ggtitle(paste0("Test data for Protein: ", test_CDK4$Protein[1]))+xlab("Drug Concentration")+ylab("Log Intensity")
```
Attached simply the script. This is (very) old code, which grew over time, thus not nice (anymore). Relevant lines are 100-139. Most of the rest you could ignore. We fit by:
try(model <- drm(response ~ dose, target, data=data, fct=modelFunction, control=drmc(method="Nelder-Mead")), silent=T)
modelFunction default is "Log-logistic (ED50 as parameter)" for us, which worked best for the kinobeads type data (in our hands), but certainly needs optimization for other data. Nelder-Mead, if I recall correctly (its some 6 years ago), helped in avoiding issues with "outlier" curves, e.g. when the last point was off or some other things.




Test curve fitting model code  ---

Defined functions needed:
```{r}
#this might not be necessary ... pretty sure Log-logistic (ED50 as parameter) == LL.4()
getModelFunctionByName <- function(modelType) {
        models <- as.data.frame(matrix(unlist(getMeanFunctions(display=F)), 22, 2, byrow=T))
        modelFunction <- eval(parse(text=paste(as.character(models[models$V2 == modelType,1]), "()", sep="")))
        return(modelFunction)
}

rsquared <- function(model) {
        meanobserved <- mean(model$origData$response)
        sstot <- sum((model$origData$response - meanobserved)^2)
        ssres <- sum(resid(model)^2)
        return(1-ssres/sstot)
}

#extract model parameters  --- mc = mean correction? 

extractInformation <- function(model, mc) {
        if(class(model) == "drc") {
                ed <- NULL
                try(ed <- ED(model, 50, display=F), silent=T)
                ed50 <- NA
                stded50 <- NA
                meanCorrection <- NA
                kd <- NA
                if (!is.null(ed)) {
                        ed50 <- ed[1]
                        stded50 <- ed[2]
			if (is.null(model$correction)) {
				#Do not use old correction anymore if new one is present
	                        meanCorrection <- mean(model$origData$correction, na.rm=T)
			} else {
				meanCorrection <- model$correction
			}
                        kd <- ed50
                        if (!is.na(meanCorrection) && meanCorrection < 1 && meanCorrection > 0) {
                                kd <- ed50 * meanCorrection
                        }
                }
                rse <- summary(model)$rseMat[1]
                r2 <- rsquared(model)
                bic <- BIC(model)
                aic <- AIC(model)
                down <- diff(model$curve[[1]](range(model$origData$dose))) < 0
                infl <- unname(coef(model)[4])
                top <- unname(coef(model)[3])
                bot <- unname(coef(model)[2])
                slope <- unname(coef(model)[1])
                return(data.frame(target=c(as.character(unique(model$origData$target))), slope=c(slope), bottom=c(bot), top=c(top), inflection=c(infl), ed50=c(ed50), stded50=c(stded50), mc=c(meanCorrection), kd=c(kd), down=c(down), r2=c(r2), rse=c(rse), bic=c(bic), aic=c(aic), clust=c(NA)))
        }
        return(data.frame(target = c(as.character(unique(model$target))), slope = c(NA), bottom = c(NA), top=c(NA), inflection=c(NA), ed50=c(NA), stded50=c(NA), mc=c(NA), kd=c(NA), down=c(NA), r2=c(NA), rse=c(NA), bic=c(NA), aic=c(NA), clust=c(NA)))
        #return(NULL)
}
```

```{r}
if(class(model) == "drc") {
                ed <- NULL
                try(ed <- ED(model, 50, display=F), silent=T)
                ed50 <- NA
                stded50 <- NA
                meanCorrection <- NA
                kd <- NA
                if (!is.null(ed)) {
                        ed50 <- ed[1]
                        stded50 <- ed[2]
			if (is.null(model$correction)) {
				#Do not use old correction anymore if new one is present
	                        meanCorrection <- mean(model$origData$correction, na.rm=T)
			} else {
				meanCorrection <- model$correction
			}
                        kd <- ed50
                        if (!is.na(meanCorrection) && meanCorrection < 1 && meanCorrection > 0) {
                                kd <- ed50 * meanCorrection
                        }
                }
                rse <- summary(model)$rseMat[1]
                r2 <- rsquared(model)
                bic <- BIC(model)
                aic <- AIC(model)
                down <- diff(model$curve[[1]](range(model$origData$dose))) < 0
                infl <- unname(coef(model)[4])
                top <- unname(coef(model)[3])
                bot <- unname(coef(model)[2])
                slope <- unname(coef(model)[1])
                return(data.frame(target=c(as.character(unique(model$origData$target))), slope=c(slope), bottom=c(bot), top=c(top), inflection=c(infl), ed50=c(ed50), stded50=c(stded50), mc=c(meanCorrection), kd=c(kd), down=c(down), r2=c(r2), rse=c(rse), bic=c(bic), aic=c(aic), clust=c(NA)))
        }
```


Test1: errors about convergence failing . . . 

"Error in optim(startVec, opfct, hessian = TRUE, method = optMethod, control = list(maxit = maxIt,  : 
  non-finite value supplied by optim
Error in drmOpt(opfct, opdfct1, startVecSc, optMethod, constrained, warnVal,  : 
  Convergence failed"

Online solution: switch inputs to vectors? 
  
```{r}
#test data CDKL
test_CDKL
test_CDK4

LogIntensities_temp<-c(test_CDKL$LogIntensities)
dose_temp<-c(test_CDKL$GROUP)
length(dose_temp)
dose<-c(1:9)
#define model function 
modelType = "Log-logistic (ED50 as parameter)"
response = "Normalized.LFQ.intensity."

modelFunction <- getModelFunctionByName(modelType)


drm(LogIntensities ~ GROUP, data=test_CDKL, fct=modelFunction)

drm(LogIntensities ~ GROUP, data=test_CDKL, fct=modelFunction, control=drmc(method="Nelder-Mead"))

doseRange <- function(data) {
        xidxs <- grep("X", colnames(data))
        x <- as.numeric(sub("X", "", colnames(data)[xidxs]))
        range(x)
}

#example from stack overflow 
fit_test <- drm(LogIntensities ~ GROUP,data=test_CDKL,  fct = LL.4())
plot(fit_test)

fit_test <- drm(LogIntensities ~ dose,data=test_CDKL,  fct = LL.4())
plot(fit_test,ylim=c(0,19))
```

Test 2 - works if dose concentrations are inputted as a numeric vector 

used LFQ intensities to fit data 

```{r}
#test 2 - works if dose concentrations are inputted as a numeric vector
length(dose_temp)
dose<-c(0,0,3,10,30,100,300,1000,3000)

test_CDKL$LogIntensities_norm<-test_CDKL$LogIntensities/18.41819
LogIntensities_norm_temp<-c(test_CDKL$LogIntensities_norm)


# 2 testings --- 
fit_test_norm <- drm(LogIntensities_norm_temp ~ dose,fct = LL.4())
summary(fit_test_norm)
plot(fit_test_norm,ylim=c(0,1))

fit_test <- drm(LogIntensities ~ dose,data=test_Q9H0U3,  fct = LL.4())
summary(fit_test)
plot(fit_test,ylim=c(0,19))

test_Q9H0U3$dose<-c(6,4,2,7,5,3,1,0,0)

	fit_test$origData <- test_CDKL
#	 fit_test$origData$correction <- test_CDKL[,paste(LogIntensities, "Correction", sep="")]



#gives same output as LL.4()
#drm(LogIntensities_temp ~ dose, fct=modelFunction)
#drm(LogIntensities_temp ~ dose, fct=modelFunction, control=drmc(method="Nelder-Mead"))

	test_Q9H0U3$dose<-c(1000,100,10,3000,300,30,3,1,1)
	
	test_Q9H0U3$dose<-log(test_Q9H0U3$dose)
a <- dr4pl(dose = test_Q9H0U3$dose,
             response = test_Q9H0U3$LogIntensities,
             method.init = "logistic")
  plot(a)
  
  
 b <- dr4pl(formula = LogIntensities~dose,
             data = test_Q9H0U3,
             method.init = "Mead",
             method.robust = "Tukey" )
  plot(b)

  #data.frame method
  c <- dr4pl(data = test_Q9H0U3,
             dose = dose,
             response = LogIntensities)
  plot(c)

  ##compatable with ggplot
  library(ggplot2) #load ggplot2
  c <- dr4pl(LogIntensities~dose,
             data = test_Q9H0U3,
             method.optim = "CG",
             trend = "decreasing" )
  d <- plot(c, x.breaks = c(.00135, .0135, .135, 1.35, 13.5))
  d + theme_grey()
```






```{r}
#define model
model<-fit_test

#test code 
if(class(model) == "drc") {
                ed <- NULL
                try(ed <- ED(model, 50, display=F), silent=T)
                ed50 <- NA
                stded50 <- NA
                meanCorrection <- NA
                kd <- NA
                if (!is.null(ed)) {
                        ed50 <- ed[1]
                        stded50 <- ed[2]
			if (is.null(model$correction)) {
				#Do not use old correction anymore if new one is present
	                        meanCorrection <- mean(model$origData$correction, na.rm=T)
			} else {
				meanCorrection <- model$correction
			}
                        kd <- ed50
                        if (!is.na(meanCorrection) && meanCorrection < 1 && meanCorrection > 0) {
                                kd <- ed50 * meanCorrection
                        }
                }
                rse <- summary(model)$rseMat[1]
                r2 <- rsquared(model)
                bic <- BIC(model)
                aic <- AIC(model)
                down <- diff(model$curve[[1]](range(model$origData$dose))) < 0
                infl <- unname(coef(model)[4])
                top <- unname(coef(model)[3])
                bot <- unname(coef(model)[2])
                slope <- unname(coef(model)[1])
               # return(data.frame(target=c(as.character(unique(model$origData$target))), slope=c(slope), bottom=c(bot), top=c(top), inflection=c(infl), ed50=c(ed50), stded50=c(stded50), mc=c(meanCorrection), kd=c(kd), down=c(down), r2=c(r2), rse=c(rse), bic=c(bic), aic=c(aic), clust=c(NA)))
        }
```





Reference code: 
```{r}
getSingleModel <- function(odata, modelType = "Log-logistic (ED50 as parameter)", response = "Normalized.LFQ.intensity.") {
	data <- odata
	data$response <- data[,response]

#	data <- data[data$dose < 3000 & data$dose != 100,]

        modelFunction <- getModelFunctionByName(modelType)
        data <- data[!is.na(data$response),]
        data <- data[is.finite(data$response),]
        model <- NULL
	if (dim(data)[1] > 0) {
	        try(model <- drm(response ~ dose, target, data=data, fct=modelFunction), silent=T)
	        if (is.null(model)) {
	                try(model <- drm(response ~ dose, target, data=data, fct=modelFunction, control=drmc(method="Nelder-Mead")), silent=T)
        	}
	}
	if(is.null(model) & response == "Normalized.LFQ.intensity.") {
		model <- getSingleModel(odata, modelType = modelType, response = "Normalized.Intensity.")
	} else {
	        if(is.null(model)) {
	                model <- data
	        } else {
			model$origData <- data
	                model$origData$correction <- data[,paste(response, "Correction", sep="")]

			#always use normalized raw correction
			#corCol <- paste("Normalized.Intensity.", "Correction", sep="")
			#if(length(grep(corCol, colnames(model$origData))) > 0) {
		        #       model$origData$correction <- data[,corCol]
			#	if (all(is.na(unique(model$origData$correction)))) {
			#		#does this make sense?
			#		model$origData$correction <- 1
			#	}
			#} else {
			#	model$origData$correction <- 1
			#}
			model$useData <- response
	        }
	}
        return(model)
}

```


~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Extra: 
drug 2 
```{r}
test_drug2<-input_files[[2]]
unique(test_drug2$Condition)
control_remove<-paste("c",1:15,sep="")


test_drug2_clean<-subset(test_drug2 ,!Condition %in% control_remove)
unique(test_drug2_clean$Condition)
QuantData2 <- dataProcess(test_drug2_clean,normalization = FALSE,MBimpute = FALSE) 
length(unique(QuantData2$ProteinLevelData$Protein))
```

