---
title: "Test from ProtienGroups.txt (using LFQ_Intensity values)"
author: "Sarah Szvetecz"
date: "2023-12-07"
output: html_document
---

Load packages
```{r}
library(MSstats)
library(readr)
library(dplyr)
library(ggplot2)
library(reshape)
library(reshape2)
```

Read combined ProteinGroups
```{r}
proteinGroups_combined <- read_tsv("/Users/sarahszvetecz/Downloads/ProteinGroups_combined.txt")

colnames(proteinGroups_combined)
View(proteinGroups_combined)
colnames(proteinGroups_combined)
columns_keep<-colnames(proteinGroups_combined)[c(1:2,175:184)]

proteinGroups_combined<-proteinGroups_combined[,colnames(proteinGroups_combined) %in% columns_keep]

unique(proteinGroups_combined$Drug)
```

Change to long format 
```{r}
long_protein_groups<-melt(proteinGroups_combined)
colnames(long_protein_groups)[2:4]<-c("Protein","Group","LFQ_Intensity")
```

look at data for proteins 
```{r}
check<-long_protein_groups %>% filter (Protein == "A0PJE2" & Group == "LFQ intensity DMSO")
var(check$log2_LFQ_Intensity)
unique(long_protein_groups$Drug) #19 

#subset results to only include values that have all 10 
temp_sub<-long_protein_groups %>% filter(Protein == protien_names[1])
nrow(temp_sub)

count_samples<-data.frame(matrix(nrow=5044,ncol=2))
colnames(count_samples)<-c("Protein","nrow")

protien_names<-unique(long_protein_groups$Protein)
length(protien_names)

#set up loop
for (i in 1:length(protien_names)){
    
temp_sub<-long_protein_groups %>% filter(Protein == protien_names[i])

count_samples$Protein[i]<-as.character(protien_names[i])
count_samples$nrow[i]<-nrow(temp_sub)
}

#subset to proteins with data for all concentration in 18 / 19 of the drugs
count_samples_sub<-count_samples %>% filter(nrow %in% c(180,190))

```


Variance calculation 
```{r}
#log2 transform values 
long_protein_groups$log2_LFQ_Intensity<-log2(long_protein_groups$LFQ_Intensity + 1)



#set up data for loop 
protien_names_sub<-unique(count_samples_sub$Protein) #using proteins with data for all drugs
groups<-unique(long_protein_groups$Group)
#protein_var_list<-data.frame(matrix(nrow=length(protien_names)*length(groups),ncol=3))

protein_var_list<-data.frame(matrix(nrow=0,ncol=3))
colnames(protein_var_list)<-c("Group","Protein","Variance")
protein_var_list_temp<-data.frame(matrix(nrow=1,ncol=3))
colnames(protein_var_list_temp)<-c("Group","Protein","Variance")


#set up loop
for (i in 1:length(protien_names_sub)){
  for ( j in 1:length(groups)){

sub_protein<-long_protein_groups %>% filter(Group == groups[j] & Protein == protien_names_sub[i])

protein_var_list_temp$Protein[1]<-as.character(protien_names_sub[i])
protein_var_list_temp$Group[1]<-as.character(groups[j])
protein_var_list_temp$Variance[1]<-var(sub_protein$log2_LFQ_Intensity)

protein_var_list<-rbind(protein_var_list,protein_var_list_temp)
}
}

#protein_var_list_firstRun<-protein_var_list


#Visualize variance 
ggplot(protein_var_list, aes(x=Group, y=Variance, color=Group)) +
  geom_boxplot() +
  labs(title="Variance distribution across all proteins",x="Concentration", y = "Variance")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
Repeat for just DMSO with filter out all 0s 
```{r}

test1<-long_protein_groups %>% filter(Group == groups[9] & Protein == protien_names_sub[1])
test1 %>% filter(log2_LFQ_Intensity != 0 )

var_list_nonzero<-data.frame(matrix(nrow=length(protien_names_sub),ncol=3))
colnames(var_list_nonzero)<-c("Protein","Variance","n_zeros")

for (i in 1:length(protien_names_sub)){
    
temp_sub1<-long_protein_groups %>% filter(Protein == protien_names_sub[i] & Group == groups[9])
temp_sub2<-temp_sub1 %>% filter(log2_LFQ_Intensity != 0 )

var_list_nonzero$Protein[i]<-as.character(protien_names_sub[i])
var_list_nonzero$Variance[i]<-var(temp_sub2$log2_LFQ_Intensity)
var_list_nonzero$n_zeros[i]<-19 -nrow(temp_sub2)

}

summary(var_list_nonzero)

#write.csv(var_list_nonzero,"final_project/derived_data/DMSO_protein_variance_nonzero.csv",row.names=F)

#Visualize variance 
ggplot(var_list_nonzero, aes(x=as.factor(n_zeros), y=Variance, color=as.factor(n_zeros))) +
  geom_boxplot( show.legend = FALSE) +
  labs(title="DMSO variance distribution stratified by numbers of zeros",x="Number of drugs with zero DMSO", y = "Variance")


```



```{r}
var_list_nonzero_sub<-var_list_nonzero[,-3]
var_list_nonzero_sub<-cbind(rep("DMSO_filterd",2841),var_list_nonzero_sub)
colnames(var_list_nonzero_sub)[1]<-"Group"

#combine values 
protein_var_list_new<-protein_var_list
protein_var_list[]
groups

protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 1000nM", "1000nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 100nM", "100nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 10nM", "10nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 30000nM", "30000nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 3000nM", "3000nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 300nM", "300nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 30nM", "30nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity 3nM", "3nM")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity DMSO", "DMSO")
protein_var_list$Group<-replace(protein_var_list$Group, protein_var_list$Group == "LFQ intensity PDPD", "PDPD")

protein_var_list_new1<-rbind(protein_var_list,var_list_nonzero_sub)


#Visualize variance 
ggplot(protein_var_list_new1, aes(x=Group, y=Variance, color=Group)) +
  geom_boxplot() +
  labs(title="Variance distribution across all proteins",x="Concentration", y = "Variance")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

summary(protein_var_list_new1[protein_var_list_new1$Group == "DMSO_filterd" , "Variance"])
```

subset to MET (P08581)
```{r}
#nomissing<-var_list_nonzero %>% filter (n_zeros == 0)
#summary(nomissing$Variance)

rnorm(1,21.91831,sqrt(29) )
rnorm(1,0,sqrt(29) )
rnorm(1,0,sqrt(2.072768) )


#subset to met & capmatinib
var_list_nonzero %>% filter (Protein == "P08581")
example_met<-long_protein_groups %>% filter (Protein == "P08581" & Drug == "Capmatinib_withCAKI")

#create simulations 

example_met$sim1<-NA;example_met$sim2<-NA;example_met$sim3<-NA;
for (i in 1:9){
example_met$sim1[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))
example_met$sim2[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))
example_met$sim3[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))

}


#create new dataset 
example_met$Group<-c("1000nM","100nM","10nM","30000nM","3000nM","300nM","30nM","3nM","DMSO")
MET_sim<- data.frame(cbind(rep(example_met$Group,3),rep(example_met$dose,3)))
MET_sim$log2_LFQ_Intensity<-NA
MET_sim$log2_LFQ_Intensity[1:9]<-example_met$sim3[1:9]
MET_sim$log2_LFQ_Intensity[10:18]<-example_met$sim1[1:9]
MET_sim$log2_LFQ_Intensity[19:27]<-example_met$sim2[1:9]

colnames(MET_sim)[1:2]<-c("Concentration","dose")
MET_sim$LogIntensities_norm<-MET_sim$log2_LFQ_Intensity/21.9183075

#visualize data 
order<-c("DMSO","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM")

ggplot(MET_sim, aes(x=factor(Concentration, level = order), y=log2_LFQ_Intensity)) +
  geom_boxplot( show.legend = FALSE) +
  labs(title="Simulated Replicates for MET", y = "log2_LFQ_Intensity")

#relative to DMSO 
ggplot(MET_sim, aes(x=factor(Concentration, level = order), y=LogIntensities_norm)) +
  geom_boxplot( show.legend = FALSE) +
  labs(title="Simulated Replicates for MET", y = "Relative Intensity")

```


Repeat original analysis --- 
original data (1 rep) = example_met
sim data (3 rep) = MET_sim

```{r}
protein_var_list %>% filter (Protein == "P08581")
example_met<-long_protein_groups %>% filter (Protein == "P08581" & Drug == "Capmatinib_withCAKI")


example_met$dose<-c(1000,100,10,30000,3000,300,30,3,1,1)

#remove PDPD
example_met<-example_met[-10,]

#relative values
example_met$LogIntensities_norm<-example_met$log2_LFQ_Intensity/21.91831


#1- drc package 
fit_test1<-drm(LogIntensities_norm ~ dose,data=example_met,  fct = LL.4())
summary(fit_test1)
plot(fit_test1,ylim=c(0,1.4))

#2 - dr4pl package
fit_test2 <- dr4pl(dose = example_met$dose,
             response = example_met$LogIntensities_norm,
             method.init = "logistic")
summary(fit_test2)
plot(fit_test2)
  

#use simulated data 
MET_sim
  
#1- drc package 
fit_test1.1<-drm(LogIntensities_norm ~ as.numeric(dose),data=MET_sim,  fct = LL.4())
summary(fit_test1.1)
plot(fit_test1.1,ylim=c(0,1.4))

#2 - dr4pl package
fit_test2.1 <- dr4pl(dose = as.numeric(MET_sim$dose),
             response = MET_sim$LogIntensities_norm,
             method.init = "logistic")
summary(fit_test2.1)
plot(fit_test2.1)
 
summary(fit_test2);summary(fit_test2.1)

#GOF - goodness of fit test only work with replicate data 
gof(fit_test2.1)
gof(fit_test2)

```
  


Created Simulated dataset for all proteins in Capmatinib_withCAKI ---
```{r}
#get list of proteins -- 
var_list_nonzero_sub.1<-var_list_nonzero_sub %>% filter(!is.na(Variance))
protein_list_sub<-unique(var_list_nonzero_sub.1$Protein)

protein_unique_drug<-long_protein_groups %>% filter ( Drug == "Capmatinib_withCAKI")
protein_unique_drug<-unique(protein_unique_drug$Protein)

protein_list_sub2<-intersect(protein_list_sub,protein_unique_drug)
length(protein_list_sub2)
# get list of drugs --- 
#drug_list<-unique(long_protein_groups$Drug)

#remove first row after loop 
final_sim_data<-example_met[1,colnames(example_met)%in% colnames(temp)]

#set up loop for all proteins all drugs 
for (i in 1:length(protein_list_sub2)){
  
  var_temp<-var_list_nonzero_sub.1[var_list_nonzero_sub.1$Protein == protein_list_sub2[1], "Variance"]
  

  #subset to protein and drug combo 
  temp<-long_protein_groups %>% filter (Protein == protein_list_sub2[i] & Drug == "Capmatinib_withCAKI")

  #create simulations 
  temp$sim1<-NA;temp$sim2<-NA;temp$sim3<-NA
  
  for (k in 1: nrow(temp)){
  
  temp$sim1[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))
  temp$sim2[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))
  temp$sim3[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))

  }
  
  final_sim_data<-rbind(final_sim_data,temp)
  }

final_sim_data<-final_sim_data[-1,]
#write.csv(final_sim_data,"final_project/derived_data/simulated_dataset_Capmatinib_withCAKI.csv",row.names=F)
```



Created Simulated dataset for all proteins and all drugs ---
```{r}
var_list_nonzero_sub

#get list of proteins -- 
var_list_nonzero_sub.1<-var_list_nonzero_sub %>% filter(!is.na(Variance))
protein_list_sub<-unique(var_list_nonzero_sub.1$Protein)

# get list of drugs --- 
drug_list<-unique(long_protein_groups$Drug)

#remove first row after loop 
final_sim_data<-example_met[1,colnames(example_met)%in% colnames(temp)]

#set up loop for all protiens all drugs 
for (i in 1:2){
  
  var_temp<-var_list_nonzero_sub.1[var_list_nonzero_sub.1$Protein == protein_list_sub[1], "Variance"]
  
  for (j in 1:length(drug_list)){

  #subset to protein and drug combo 
  temp<-long_protein_groups %>% filter (Protein == protein_list_sub[i] & Drug == drug_list[j])

  #create simulations 
  temp$sim1<-NA;temp$sim2<-NA;temp$sim3<-NA
  
  for (k in 1: nrow(temp)){
  
  temp$sim1[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))
  temp$sim2[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))
  temp$sim3[k]<-rnorm(1,temp$log2_LFQ_Intensity[k],sqrt(var_temp))

  }
  
  final_sim_data<-rbind(final_sim_data,temp)
  }
}

```

  
(same as of above with different protein) 
Repeat original analysis (can insert different protein names here to test)--- 
original data (1 rep) = example_ATP
sim data (3 rep) = MET_sim

subset to ATP6V1H (Q9UI12)
```{r}
#calculate variance from no missing 0s
nomissing<-var_list_nonzero %>% filter (n_zeros == 0)
summary(nomissing$Variance)

#test
rnorm(1,21.91831,sqrt(29) )
rnorm(1,0,sqrt(29) )
rnorm(1,0,sqrt(2.072768) )


var_list_nonzero %>% filter (Protein == "Q9UI12")
example_ATP<-long_protein_groups %>% filter (Protein == "Q9UI12" & Drug == "Capmatinib_withCAKI")

#create simulations 

example_met$sim1<-NA;example_met$sim2<-NA;example_met$sim3<-NA;
for (i in 1:9){
example_met$sim1[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))
example_met$sim2[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))
example_met$sim3[i]<-rnorm(1,example_met$log2_LFQ_Intensity[i],sqrt(2.072768))

}


#create new dataset 
example_met$Group<-c("1000nM","100nM","10nM","30000nM","3000nM","300nM","30nM","3nM","DMSO")
MET_sim<- data.frame(cbind(rep(example_met$Group,3),rep(example_met$dose,3)))
MET_sim$log2_LFQ_Intensity<-NA
MET_sim$log2_LFQ_Intensity[1:9]<-example_met$sim3[1:9]
MET_sim$log2_LFQ_Intensity[10:18]<-example_met$sim1[1:9]
MET_sim$log2_LFQ_Intensity[19:27]<-example_met$sim2[1:9]

colnames(MET_sim)[1:2]<-c("Concentration","dose")
MET_sim$LogIntensities_norm<-MET_sim$log2_LFQ_Intensity/21.9183075

#visualize data 
order<-c("DMSO","3nM","10nM","30nM","100nM","300nM","1000nM","3000nM","30000nM")

ggplot(MET_sim, aes(x=factor(Concentration, level = order), y=log2_LFQ_Intensity)) +
  geom_boxplot( show.legend = FALSE) +
  labs(title="Simulated Replicates for MET", y = "log2_LFQ_Intensity")

#relative to DMSO 
ggplot(MET_sim, aes(x=factor(Concentration, level = order), y=LogIntensities_norm)) +
  geom_boxplot( show.legend = FALSE) +
  labs(title="Simulated Replicates for MET", y = "Relative Intensity")

```


EXTRA scratch code ------ 


Subset to MET (P08581)
```{r}
protein_var_list %>% filter (Protein == "P08581")
example_met<-long_protein_groups %>% filter (Protein == "P08581" & Drug == "Capmatinib_withCAKI")


example_met$dose<-c(1000,100,10,30000,3000,300,30,3,1,1)

#remove PDPD
example_met<-example_met[-10,]

#relative values
example_met$LogIntensities_norm<-example_met$log2_LFQ_Intensity/21.91831


#1- drc package 
fit_test1<-drm(LogIntensities_norm ~ dose,data=example_met,  fct = LL.4())
summary(fit_test1)
plot(fit_test1,ylim=c(0,1.4))

#2 - dr4pl package
fit_test2 <- dr4pl(dose = example_met$dose,
             response = example_met$LogIntensities_norm,
             method.init = "logistic")
summary(fit_test2)
plot(fit_test2)
  

#use simulated data 
MET_sim
  
#1- drc package 
fit_test1.1<-drm(LogIntensities_norm ~ as.numeric(dose),data=MET_sim,  fct = LL.4())
summary(fit_test1.1)
plot(fit_test1.1,ylim=c(0,1.4))

#2 - dr4pl package
fit_test2.1 <- dr4pl(dose = as.numeric(MET_sim$dose),
             response = MET_sim$LogIntensities_norm,
             method.init = "logistic")
summary(fit_test2.1)
plot(fit_test2.1)
 
summary(fit_test2);summary(fit_test2.1)

gof(fit_test2.1)
gof(fit_test2)
```


Other tests 
```{r}
 b <- dr4pl(formula = LogIntensities_norm~dose,
             data = example_met,
             method.init = "Mead",
             method.robust = "Tukey" )
  plot(b)

  #data.frame method
  c <- dr4pl(data = example_met,
             dose = dose,
             response = LogIntensities_norm)
 summary(c)
   plot(c)

  ##compatable with ggplot
  library(ggplot2) #load ggplot2
  c <- dr4pl(LogIntensities_norm~dose,
             data = example_met,
             method.optim = "CG",
             trend = "decreasing" )
  d <- plot(c, x.breaks = c(.00135, .0135, .135, 1.35, 13.5))
  d + theme_grey()
```



```{r}
#test 2 - works if dose concentrations are inputted as a numeric vector
length(dose_temp)
dose<-c(0,0,3,10,30,100,300,1000,3000)

test_CDKL$LogIntensities_norm<-test_CDKL$LogIntensities/18.41819
LogIntensities_norm_temp<-c(test_CDKL$LogIntensities_norm)


# 2 testings --- 
fit_test_norm <- drm(LogIntensities_norm_temp ~ dose,fct = LL.4())
summary(fit_test_norm)
plot(fit_test_norm,ylim=c(0,1))

fit_test <- drm(LogIntensities ~ dose,data=test_Q9H0U3,  fct = LL.4())
summary(fit_test)
plot(fit_test,ylim=c(0,19))

test_Q9H0U3$dose<-c(6,4,2,7,5,3,1,0,0)

	fit_test$origData <- test_CDKL
#	 fit_test$origData$correction <- test_CDKL[,paste(LogIntensities, "Correction", sep="")]



#gives same output as LL.4()
#drm(LogIntensities_temp ~ dose, fct=modelFunction)
#drm(LogIntensities_temp ~ dose, fct=modelFunction, control=drmc(method="Nelder-Mead"))

	test_Q9H0U3$dose<-c(1000,100,10,3000,300,30,3,1,1)
	
	test_Q9H0U3$dose<-log(test_Q9H0U3$dose)
a <- dr4pl(dose = test_Q9H0U3$dose,
             response = test_Q9H0U3$LogIntensities,
             method.init = "logistic")
  plot(a)
  
  
 b <- dr4pl(formula = LogIntensities~dose,
             data = test_Q9H0U3,
             method.init = "Mead",
             method.robust = "Tukey" )
  plot(b)

  #data.frame method
  c <- dr4pl(data = test_Q9H0U3,
             dose = dose,
             response = LogIntensities)
  plot(c)

  ##compatable with ggplot
  library(ggplot2) #load ggplot2
  c <- dr4pl(LogIntensities~dose,
             data = test_Q9H0U3,
             method.optim = "CG",
             trend = "decreasing" )
  d <- plot(c, x.breaks = c(.00135, .0135, .135, 1.35, 13.5))
  d + theme_grey()
```


Subset to ITG1 
```{r}
protein_var_list %>% filter (Protein == "P05556")
long_protein_groups %>% filter (Protein == "P05556" & Drug == "Capmatinib_withCAKI")

#PRDX4
protein_var_list %>% filter (Protein == "Q13162")
example_noeffect<-long_protein_groups %>% filter (Protein == "Q13162" & Drug == "Capmatinib_withCAKI")

```

```{r}
example_noeffect$dose<-c(1000,100,10,30000,3000,300,30,3,1,1)

#remove PDPD
example_noeffect<-example_noeffect[-10,]

#relative values
example_noeffect$LogIntensities_norm<-example_noeffect$log2_LFQ_Intensity/22.67418


#1- drc package 
fit_test1<-drm(LogIntensities_norm ~ dose,data=example_noeffect,  fct = LL.4())
summary(fit_test1)
plot(fit_test1,ylim=c(0,1.4))
```

